Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-10-31 manuscript/manuscript.qmd
Progress with introduction and new models

targets::tar_load(
	c(participants,
	  stimuli,
	  vocabulary,
	  attrition_trials, 
	  attrition_participants,
	  bvq_data, 
	  gaze_aoi, 
	  data_related,
	  data_cognate,
	  data_time_related, 
	  data_time_cognate,
	))

targets::tar_load(
	c(model_fits_aggr,
	  model_fits_time,
	  model_loos_aggr,
	  model_loos_time,
	))

@@

Growing a mental lexicon is one of the main achievements in the developmental journey of infants. Being able to store lexical representations of familiar words allows infants to later recognise them in speech, and to retrieve their associated concepts. The first step towards spoken word-recognition is the acoustic-phonetic processing of the speech signal. This results in the activation of lexical representations whose associated phonological forms matches the acoustic signal to some extent. The activated word-forms then compete for selection: the lexical system uses the auditory input, together with top-down context provided by grammatical, supra-segmental, and semantic constraints, to select the candidate lexical representation with the best match. There is evidence that spoken word recognition occurs in a *cascaded* fashion. The activation and selection of candidate lexical representations overlap in time, relying on partial information that accumulates as the speech signal unfolds [@marslen1987functional; @marslen1978processing; @grosjean1980spoken]. This cascaded activation triggered by lexical access also results in the activation of the concepts associated to the accesses phonological forms. @allopenna1998tracking.

This is also the case in the developing lexicon. There is evidence that infants start forming their first lexical representations before their first year of life. It is well established that, across languages, infants start recognising their first word-forms before their first year of life [@jusczyk1999; @halle1994; @vihman2004]. Already at six months of age, infants show evidence of word recognition. When presented with pairs of familiar objects and hearing the label of one of them, 6-months-old infants and older show a looking preference towards the named object [@bergelson20126; @bergelson2015early]. @chow2017spoken adapted the Visual World Paradigm from @huettig2007tug to explore 24- to 30-month-old toddlers' visual fixation patterns during a word recognition when presented with phonological and semantic distractors. In each trial, the authors presented participants with four semantically and phonologically pictures. Four seconds after pictures onset, a word-form was auditorily presented. The word did not refer to any of the pictures displayed on the screen, but was phonologically related to one of them (both labels shared phonological onset), and semantically related to another one of the pictures (both referents belonged to the same taxonomic category. For instance, participants might be presented with the pictures of a sandwich, a bus, a cat, and a dress. Then they would hear the carrier phrase "Look at the **bee**!". The authors registered participants fixations to the phonological and semantic distractors, and found evidence of a preference for the phonological distractor at earlier stages of the post-naming phase, and a preference for the semantic distractor at later stages of the trial. These results support a cascaded activation account of lexical access during the first stages of lexical development.

Evidence from cascaded activation in lexical access is also provided by bilinguals.A critical property of bilingual language processing, both in adults and infants, is that of non-selective access to the lexicon. When listening to speech, bilinguals activate all lexical candidates from either language consistent with the speech signal. Even more surprising, when speaking, bilinguals activate all potential words consistent with the concepts they want to express.
This effect has been systematically reported in adults [@marian1999activation; @costa2000cognate; @thierry2007brain; @dunabeitia2009masked].

Bilinguals infants start recognising words at similar ages. Using a Head-turn Preference Procedure and ERP registration,  @vihman2013 found English-Welsh bilinguals to be sensitive to the prevalence of auditorily presented words. Infants' looking times and ERP differed when presented with words frequently marked as "familiar" by caregivers of demographically equivalent peers (e.g., *nappy*, *thank you*, *bottle*, *apple*, and *naughty*), compared to when presented with rare words (e.g., *Eiffel*, *courtly*, *nettle*, *wacky*, and *juncture*). Bilinguals' trajectories of word recognition  [@vihman2013].

Bilingual infants receive a relative impoverished linguistic input in each of their languages compared to their monolingual peers. There is not reason to think bilinguals receive an increased amount of total linguistic input, so the fact that such input is divided between at least two languages means that they receive less input from each of their languages [@costa2014does]. The input bilinguals receive is also more complex than that of monolinguals. This is reflected in multiple dimensions. For instance, bilinguals are presented with two different, but possible overlapping phoneme inventories, they may be presented with two different grammatical systems, and often presented with more than one label per referent. This does not keep bilinguals from tuning their perceptual abilities to both phoneme inventories of their native languages, distinguishing grammatical irregularities in both languages, or learning more than one label per object. Bilinguals' developmental trajectories might not always be identical to monolinguals', but they achieve their language acquisition milestones at equivalent ages.

For instance, previous work on lexical acquisition has provided some insight into how bilinguals' vocabulary size is equivalent to monolinguals' across ages, and the possible mechanisms behind this. @hoff2012 found that English-Spanish bilinguals' vocabulary in English was influenced by the words they had already acquired in Spanish. This is consistent with previous accounts of lexical acquisition in which infants acquire new words through preferential acquisition [@hill2013, @fourtassi2020]. A monograph by @floccia2018introduction provided evidence of a language distance effect, where bilinguals learning two languages sharing higher lexical similarity knew more words in their less dominant language than those learning two language with lower lexical similarity. Later work by @mitchell2022cognate and @garcia-castro@cognate provided evidence for a more direct facilitation effect of cognateness. Cognates are form-similar translation equivalents. Cognateness is suggested to facilitate word acquisition through parallel activation: the co-activation of translation equivalents in speech might lead to an faster accumulation of learning instances with words, which ultimately results in an earlier age of acquisition.

@@

@@

## Evidence from co-activation though translation
The role of cognates in lexical processing.

@@

## Implicit naming

The aforementioned studies use priming-though-translation paradigms. In these paradigms, cross-linguistic priming results from activation spreading from the lexical representation of the the (auditorily presented) prime word to the lexical representation of its translation through semantic associations (e.g., translation equivalence). When prime words are auditorily presented, the task design constraints the naming context to one of the languages. The consequence is a potential asymmtry between the activation dynamics of the lexical represetnation of both labels: while the participant has heard one label, the activation of its translation must spread across semantic links. This might lead to a lower strength, and higher lacenty of the activation of the translation.

##  How does language co-activation shape lexical development?

@@

@@

Missing eye-tracker samples were interpolated using the last-observation-carried-forward [see @zettersten2022peekbank for a similar approach], with a maximum of 20 maximum consecutive missing samples being interpolated (an equivalent of 166.67 ms).

@@

```{r fig-ptlt-descriptive}
#| label: fig-ptlt-descriptive
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.elog = mean(.elog),
			  .by = c(session_id, timebin, lp, condition)) |> 
	ggplot(aes(timebin, .elog, colour = condition, fill = condition)) +
	facet_wrap(~lp) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = mean_se,
				 geom = "errorbar",
				 width = 0.5) +
	stat_summary(fun = mean,
				 geom = "line") +
	stat_summary(fun = mean,
				 geom = "point") +
	theme(axis.title.x = element_blank()) +

	data_time |> 
	summarise(.elog = mean(.elog),
			  .by = c(session_id, timebin, lp, condition)) |> 
	ggplot(aes(timebin, .elog, colour = condition, fill = condition)) +
	facet_wrap(~lp) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = mean_se,
				 geom = "errorbar",
				 width = 0.5) +
	stat_summary(fun = mean,
				 geom = "line") +
	stat_summary(fun = mean,
				 geom = "point") +
	theme(strip.text = element_blank()) +

	plot_layout(ncol = 1) &
	plot_annotation(tag_levels = "A") &
	labs(x = "Time bin (100 ms)",
		 y = "P(Target looking)",
		 colour = "Condition",
		 fill = "Condition") &
	# scale_y_continuous(breaks = seq(0, 1, 0.1)) &
	theme(legend.position = "right",
		  panel.grid.major.x = element_line(colour = "grey",
		  								  linetype = "dotted"),
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))

@@

```{r fig-related-aggregated}
#| label: fig-related-aggregated
#| fig-cap: "A) Aggregated probability of target looking for related and unrelated trials. Each point represents the mean probability of target looking for one participant in one testing session. Lines link observations from the same testing session across conditions. Black dots and errorbars indicate the mean and 95% CrI. B) Aggregated probability of target looking across ages for related and unrelated trials. Lines and error bands indicate the mean and standard error of the mean."
nd <- expand(data_aggr, lp, condition, age = seq(20, 32, length.out = 100))

epreds <- add_epred_draws(newdata = nd,
						  object = model_fits_aggr$fit_aggr_4,
						  re_formula = NA) |> 
	mean_qi() |> 
	mutate(label = sprintf("%.2f\n[%.2f, %.2f]", .epred, .lower, .upper))

epreds |> 
	ggplot(aes(age, .epred, 
			   colour = condition,
			   fill = condition)) +
	facet_wrap(~lp) +
	# geom_line(data = data_related,
	# 		  aes(y = .elog,
	# 		  	group = session_id),
	# 		  colour = "grey",
	# 		  alpha = 1/4) +
	# geom_dots(dat = data_aggr,
	# 		  aes(y = .elog),
	# 		  side = "both",
	# 		  show.legend = FALSE) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	geom_ribbon(aes(ymin = .lower,
					ymax = .upper),
				alpha = 1/3,
				linewidth = NA) +
	geom_line(linewidth = 3/4) +
	geom_point(data = data_aggr,
			   aes(y = .elog),
			   alpha = 1/2) +
	# geom_label(aes(label = label,
	# 			   y = -1.5),
	# 		   fill = "white",
	# 		   colour = "black",
	# 		   label.r = unit(0, "lines"),
	# 		   label.size = 0,
	# 		   vjust = 0,
	# 		   size = 3,
	# 		   alpha = 1/2) +
	labs(x = "Condition",
		 y = "P(Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_y_continuous(breaks = seq(-4, 4, 0.5)) +
	theme(legend.position = "none",
		  axis.title.x = element_blank(),
		  strip.background = element_rect(fill = "grey90",
		  								colour = "grey90"),
		  panel.grid.major.y = element_line(linetype = "dotted",
		  								  colour = "grey")) 
```

@@

```{r fig-time}
#| label: fig-time
#| fig-height: 6
#| fig-width: 8
#| fig-cap: "Marginal posterior predictions of the GAMMs in Analysis 1. (A) Mean posterior probability of target looking across the time course of the trial. Black lines and intervals indicate the psoterior mean and 95% credible intervals. Points indicate the mean probability of target looking across participants. (B) Difference in posterior probability of target looking between *related* and *unrelated* trials. The yellow rectangle indicates, in both A and B, the range of time points in which the 95% credible interval of the differences excluded zero."
plot_gamm <- function(data, model, contrast = "related") {

	nd <- expand_grid(condition = levels(data$condition),
					  timebin = seq(0, 17, length.out = 100),
					  age = mean(data$age),
					  # sesion_id = first(data$session_id),
					  # age = c(21, 25, 30),
					  lp = levels(data$lp))

	epreds <- add_epred_draws(newdata = nd,
							  object = model,
							  # ndraws = NULL,
							  re_formula = NA,
							  value = ".value") |> 
		mutate(lp = factor(lp, 
						   levels = c("Monolingual (English)",
						   		   "Monolingual",
						   		   "Bilingual")),
			   age = factor(age, levels = c(21, 25, 30),
			   			 labels = paste0(c(21, 25, 30), " months")))

	epreds_diff <- epreds |> 
		pivot_wider(names_from = condition,
					values_from = .value,
					id_cols = c(timebin, age, lp, .draw),
					names_repair = janitor::make_clean_names) 
	# 
	# if (contrast=="related") {
	# 	epreds_diff$diff <- epreds_diff$related - epreds_diff$unrelated
	# } else {
	# 	epreds_diff$diff <- epreds_diff$cognate - epreds_diff$non_cognate
	# }

	obs <- data |> 
		mutate(age = factor(age, levels = c(21, 25, 30),
							labels = paste0(c(21, 25, 30), " months"))) |> 
		summarise(.elog = mean(.elog),
				  .by = c(session_id, timebin, lp, condition, age))

	epreds |> 
		ggplot(aes(timebin, y = .value,
				   colour = condition,
				   fill = condition,
				   shape = condition)) +
		facet_wrap(~lp) +
		# stat_summary(fun.data = mean_qi,
		# 			 geom = "ribbon",
		# 			 alpha = 0.5,
		# 			 linewidth = 0) +
		stat_summary(aes(linetype = condition),
					 fun = mean,
					 geom = "line",
					 linewidth = 3/4) +
		geom_hline(yintercept = 0, 
				   linewidth = 1/2,
				   colour = "black",
				   linetype = "dotted") +
		stat_summary(data = obs,
					 aes(y = .elog),
					 fun.data = "mean_se",
					 geom = "errorbar",
					 width = 1,
					 linewidth = 3/4,
					 position = position_dodge(width = 0.5)) +
		stat_summary(data = obs,
					 aes(y = .elog),
					 fun = "mean",
					 geom = "point",
					 size = 2,
					 position = position_dodge(width = 0.5)) +
		# stat_summary(fun.data = mean_se,
		# 			 geom = "errorbar",
		# 			 linetype = "solid",
		# 			 width = 0.25) +
		# stat_summary(fun = mean,
		# 			 geom = "point") +
		labs(x = "Time (ms)",
			 y = "P(Target looking)",
			 colour = "Condition",
			 fill = "Condition",
			 linetype = "Condition",
			 shape = "Condition") +
		theme(legend.title = element_blank(),
			  legend.position = "top",
			  axis.title.x = element_blank())

	epreds_diff |> 
		ggplot(aes(timebin, diff)) +
		facet_wrap(~lp) +
		stat_lineribbon(.width = 0.95,
						linewidth = 0,
						fill = "grey") +
		stat_summary(data = epreds_diff,
					 fun = "mean",
					 geom = "line",
					 colour = "black",
					 linewidth = 3/4) +
		geom_hline(yintercept = 0, 
				   linewidth = 1/2,
				   colour = "black",
				   linetype = "dotted") +
		labs(x = "Time (ms)",
			 y = "Related - Unrelated",
			 fill = "CrI") +
		theme(strip.text = element_blank(),
			  legend.position = "none") +

		plot_layout(ncol = 1) &
		plot_annotation(tag_levels = "A") +
		scale_linetype_manual(values = rev(c("solid", "dashed"))) &
		# scale_shape_manual(values = c(1, 2)) &
		scale_x_continuous(labels = \(x) format((x * 1e2), 
												big.mark = ",")) &
		theme(panel.grid = element_blank(),
			  legend.position = "top") 
}

plot_gamm(data_time_related, 
		  model_fits_related[[1]],
		  contrast = "related")

>>>>

targets::tar_load(
	c(participants,
	  stimuli,
	  vocabulary,
	  attrition_trials, 
	  attrition_participants,
	  bvq_data, 
	  gaze_aoi, 
	  data_time, 
	  data_aggr,
	))

targets::tar_load(
	c(model_fits_aggr,
	  model_fits_time
	  # model_loos_aggr,
	  # model_loos_time,
	))

@@

Building a mental lexicon is a major achievement in the developmental journey of infants: by accumulating lexical representations of familiar words, infants are later able to recognise them in speech, and to extract meaning from the speech they are exposed to. Infants start forming their first lexical representations before their first year of life [@jusczyk1999; @halle1994; @vihman2004]. When presented with pairs of pictures of familiar objects, 6-months-old infants already direct their gaze towards whichever is named [@bergelson20126; @bergelson2015early]. This initial lexicon is hardly larger than around fifteen words [@REFERENCE], but it undergoes rapid growth [@REFERENCE] during their second year of life, at the end of which toddlers know an average of XXX words, and by 30 months of age, they know XXX words [@REFERENCE]. This accelerated lexical developmental during the second year of life is also reflected in infants'  trajectories of word recognition: infants recognise familiar words faster and more efficiently [@fernald1998rapid; @fernald2001half; @hurtado2007spoken]. Bilingual infants' trajectories of lexical acquisition [@hoff2012, @de2014bilingual; @pearson1994patterns; @byers-heinlein2023sometimes] and word recognition [@vihman2007onset; @vihman2013; @legacy2018vocabulary] are equivalent to their monolingual peers'. This is a remarkable deed: bilingual infants receive a relative impoverished linguistic input in each of their languages compared to monolinguals [@costa2014does]. The mechanism that facilitate bilingual lexical development are still unclear.

## (Early) Word recognition is cascaded

Despite their different linguistic input, monolinguals and bilinguals share broadly common word recognition mechanisms. The first step towards spoken word-recognition is the acoustic-phonetic processing of the speech signal. This results in the activation of lexical representations whose associated phonological forms matches the acoustic signal to some extent. The activated word-forms then compete for selection: the lexical system uses the auditory input, together with top-down context provided by grammatical, supra-segmental, and semantic constraints, to select the candidate lexical representation with the best match. There is evidence that spoken word recognition occurs in a *cascaded* fashion. The activation and selection of candidate lexical representations overlap in time, relying on partial information that accumulates as the speech signal unfolds [@marslen1987functional; @marslen1978processing; @grosjean1980spoken]. This cascaded activation triggered by lexical access also results in the activation of the concepts associated to the accesses phonological forms. @allopenna1998tracking.

This is also the case of the developing lexicon: like adults, infants activate non-selected lexical representations. This is relfected by the fact that from 21 months of age, infants' lexicon seems to be systematically organised accordi to the semantic and phnological properties of the lexical representations.

SEMANTICS
@arias2009lexical showed evidence of semantic priming in 21-month-old infants. In a word recognition task, infants' recognition of a spoken word (e.g., *dot*) was interfered by the previous presentation of a semantically related word (e.g., *cat*). Later studies provided electrophysiological evidence of such effect as early as 14 months of age [@friedrich2005lexical].

PHONOLOGY
Initial lexical representations are also phonologically well-specified [@swingley2000spoken; @swingley2002lexical]. Overall these results suggest that early lexical representations are systematically organised in the lexicon according to their semantic and phonological properties.

BOTH
@chow2017spoken adapted the Visual World Paradigm from @huettig2007tug to explore 24- to 30-month-old toddlers' visual fixation patterns during a word recognition when presented with phonological and semantic distractors. In each trial, the authors presented participants with four semantically and phonologically pictures. Four seconds after pictures onset, a word-form was auditorily presented. The word did not refer to any of the pictures displayed on the screen, but was phonologically related to one of them (both labels shared phonological onset), and semantically related to another one of the pictures (both referents belonged to the same taxonomic category. For instance, participants might be presented with the pictures of a sandwich, a bus, a cat, and a dress. Then they would hear the carrier phrase "Look at the **bee**!". The authors registered participants fixations to the phonological and semantic distractors, and found evidence of a preference for the phonological distractor at earlier stages of the post-naming phase, and a preference for the semantic distractor at later stages of the trial. These results support a cascaded activation account of lexical access during the first stages of lexical development. @swingley1999continuous

## Cascaded activation in bilingual parallel activation

There is also evidence of cascaded activation in bilinguals. A critical property of bilingual language processing, both in adults and infants, is that of non-selective access to the lexicon. When listening to speech, bilinguals activate all lexical candidates from either language consistent with the speech signal. Even more surprising, when speaking, bilinguals activate all potential words consistent with the concepts they want to express. This effect has been systematically reported in adults [@marian1999activation; @costa2000cognate; @thierry2007brain; @dunabeitia2009masked].

For instance, previous work on lexical acquisition has provided some insight into how  found that English-Spanish bilinguals' vocabulary in English was influenced by the words they had already acquired in Spanish. This is consistent with previous accounts of lexical acquisition in which infants acquire new words through preferential acquisition [@hill2013, @fourtassi2020].

A monograph by @floccia2018introduction provided evidence of a language distance effect, where bilinguals learning two languages sharing higher lexical similarity knew more words in their less dominant language than those learning two language with lower lexical similarity. Later work by @mitchell2022cognate and @garcia-castro@cognate provided evidence for a more direct facilitation effect of cognateness. Cognates are form-similar translation equivalents. Cognateness is suggested to facilitate word acquisition through parallel activation: the co-activation of translation equivalents in speech might lead to an faster accumulation of learning instances with words, which ultimately results in an earlier age of acquisition.

@@

## Cognate word recogntiion

@@

@@

## The rpesent study

The aforementioned studies use priming-though-translation paradigms. In these paradigms, cross-linguistic priming results from activation spreading from the lexical representation of the the (auditorily presented) prime word to the lexical representation of its translation through semantic associations (e.g., translation equivalence). When prime words are auditorily presented, the task design constraints the naming context to one of the languages. The consequence is a potential asymmtry between the activation dynamics of the lexical representation of both labels: while the participant has heard one label, the activation of its translation must spread across semantic links. This might lead to a lower strength, and higher lacenty of the activation of the translation.

@@

We compared the explanatory power of three measures of vocabulary size on participants' trajectories of word recognition: L1 vocabulary and total vocabulary.Previous studies suggest that same-language vocabulary size predicts better bilingual infants' performance in word recognition tasks, as compared to cross-language vocabulary size [@marchman2010vocabulary]

@@

We defined a time windows of interest from 200 ms after target and distractor pictures onset until the end of the trial at 2,000 ms when  both pictures disappeared from screen. To avoid modelling fixations driven by processes other than auditory word recognition [@fernald1998rapid; @fernald2001half]. Missing eye-tracker samples were interpolated using the last-observation-carried-forward [see @zettersten2022peekbank for a similar approach], with a maximum of 20 maximum consecutive missing samples being interpolated (an equivalent of 166.67 ms).

@@

## Preliminary analyses

```{r fig-ptlt-recognition}
#| label: fig-ptlt-recognition-aggr
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = lp)) +
	facet_wrap(~lp, nrow = 1, scales = "free_x") +
	geom_line(aes(group = session_id),
			  # colour = "grey",
			  alpha = 1/10) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT") +
	scale_x_continuous(breaks = seq(0, 20, 2),
					   labels = \(x) format(x*100+300, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.2),
					   limits = c(0, 1)) +
	theme(panel.grid.major.y = element_line(colour = "grey",
											linetype = "dotted"),
		  legend.position = "none")
```

```{r fig-ptlt-recognition-age}
#| label: fig-ptlt-recognition-age
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, age_group, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = age_group)) +
	facet_wrap(~lp, nrow = 1, scales = "free_x") +
	# geom_rug(aes(x = timebin[is.na(.prop)])) +
	# geom_line(aes(group = session_id),
	# 		  alpha = 1/8) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT",
		 colour = "Age") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.05)) +
	theme(legend.position = "top",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))
```

```{r fig-ptlt-recognition-condition}
#| label: fig-ptlt-recognition-condition
#| fig-cap: "Average probability of target looking across the time course of the trial. Points an whiskers indicate the mean and standard error of the mean."
data_time |> 
	summarise(.prop = mean(.prop),
			  .by = c(lp, timebin, condition, session_id)) |> 
	ggplot(aes(timebin, .prop, colour = condition)) +
	facet_grid(~lp) +
	# geom_rug(aes(x = timebin[is.na(.prop)])) +
	# geom_line(aes(group = interaction(condition, session_id)),
	# 		  alpha = 1/8) +
	geom_hline(yintercept = 0.5, linetype = "dashed") +
	stat_summary(fun.data = "mean_se",
				 geom = "errorbar",
				 width = 0.35,
				 linewidth = 2/4) +
	stat_summary(fun = "mean",
				 geom = "point",
				 size = 1.2) +
	stat_summary(fun = "mean",
				 geom = "line",
				 linewidth = 2/4) +
	labs(x = "Time (ms)",
		 y = "PTLT",
		 colour = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	scale_y_continuous(labels = scales::percent,
					   breaks = seq(0, 1, 0.05)) +
	theme(legend.position = "top",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))
```

@@

```{r fig-related-aggregated}
#| label: fig-related-aggregated
#| fig-cap: "A) Aggregated probability of target looking for related and unrelated trials. Each point represents the mean probability of target looking for one participant in one testing session. Lines link observations from the same testing session across conditions. Black dots and errorbars indicate the mean and 95% CrI. B) Aggregated probability of target looking across ages for related and unrelated trials. Lines and error bands indicate the mean and standard error of the mean."
plot_agg_epreds <- function(newdata, object) {

	epreds <- add_epred_draws(newdata = newdata,
							  object = object,
							  re_formula = NA)
	return(mean_qi(epreds))
}

voc_values <- c(0.25, 0.5, 0.75, 1.00)
age_values <- c(21, 25, 30)

nd_0 <- expand(data_aggr, lp, condition, age = age_values)
nd_1 <- expand(data_aggr, lp, condition, voc_l1 = voc_values)
nd_2 <- expand(data_aggr, lp, condition, voc_total = voc_values)

epreds <- map2(.x = list(nd_0, nd_1, nd_2),
			   .y = model_fits_aggr,
			   .f = plot_agg_epreds) |> 
	map(\(x) {
		mutate(x, 
			   across(matches("voc"),
			   	   \(x) factor(x = x, 
			   	   			levels = voc_values,
			   	   			labels = scales::percent(voc_values))),
			   across(matches("age"),
			   	   \(x) factor(x = x,
			   	   			levels = age_values,
			   	   			labels = age_values)))
	}) |> 
	map(\(x) rename_with(x, \(x) gsub("age|voc_total|voc_l1", "var", x))) |> 
	set_names(names(model_fits_aggr))

plot_epreds <- function(data, x) {
	data |> 
		ggplot(aes(var, .epred, 
				   colour = condition,
				   fill = condition)) +
		facet_wrap(~lp, scales = "free_x") +
		geom_hline(yintercept = 0,
				   linetype = "dashed") +
		geom_errorbar(aes(ymin = .lower,
						  ymax = .upper),
					  width = 0.5,
					  linewidth = 3/4,
					  position = position_dodge(width = 0.5)) +
		geom_point(position = position_dodge(width = 0.5),
				   size = 2)
}

plot_epreds(epreds$fit_aggr_0) +
	labs(x = "Age (months)") +
	plot_epreds(epreds$fit_aggr_1) +
	labs(x = "L1 vocabulary size") +
	theme(strip.text = element_blank()) +
	plot_epreds(epreds$fit_aggr_2) +
	theme(strip.text = element_blank()) +
	labs(x = "Total vocabulary size") +
	plot_layout(ncol = 1, guides = "collect") &
	labs(y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") &
	theme(legend.position = "top",
		  panel.grid.major.y = element_line(colour = "grey",
		  								  linetype = "dotted"))

@@

```{r fig-time}
#| label: fig-time
#| fig-height: 6
#| fig-width: 8
#| fig-cap: "Marginal posterior predictions of the GAMMs in Analysis 1. (A) Mean posterior probability of target looking across the time course of the trial. Black lines and intervals indicate the psoterior mean and 95% credible intervals. Points indicate the mean probability of target looking across participants. (B) Difference in posterior probability of target looking between *related* and *unrelated* trials. The yellow rectangle indicates, in both A and B, the range of time points in which the 95% credible interval of the differences excluded zero."
nd <- expand_grid(condition = levels(model_fits_time$fit_time_2$data$condition),
				  timebin = seq(0, 17, length.out = 100),
				  voc_total = mean(model_fits_time$fit_time_2$data$voc_total),
				  # sesion_id = first(data$session_id),
				  # age = c(21, 25, 30),
				  lp = levels(model_fits_time$fit_time_2$data$lp))

epreds <- add_epred_draws(newdata = nd,
						  object = model_fits_time$fit_time_2,
						  ndraws = 20,
						  re_formula = NA,
						  value = ".value") |> 
	mean_qi() |> 
	mutate(lp = factor(lp, 
					   levels = c("Monolingual (English)",
					   		   "Monolingual",
					   		   "Bilingual")))

epreds |> 
	ggplot(aes(timebin, .value,
			   colour = condition,
			   fill = condition)) +
	facet_wrap(~lp) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	geom_ribbon(aes(ymin = .lower, ymax = .upper),
				alpha = 0.5,
				linewidth = 0) +
	geom_line(linewidth = 3/4) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	theme(legend.position = "top",
		  panel.grid.major = element_line(colour = "grey",
		  								linetype = "dotted"))
```

```{r fig-time-total}
#| label: fig-time-total
#| fig-height: 6
#| fig-width: 8
#| fig-cap: "Marginal posterior predictions of the GAMMs in Analysis 1. (A) Mean posterior probability of target looking across the time course of the trial. Black lines and intervals indicate the psoterior mean and 95% credible intervals. Points indicate the mean probability of target looking across participants. (B) Difference in posterior probability of target looking between *related* and *unrelated* trials. The yellow rectangle indicates, in both A and B, the range of time points in which the 95% credible interval of the differences excluded zero."
voc_values <- seq(0.25, 1, 0.25)

nd <- expand_grid(condition = levels(model_fits_time$fit_time_2$data$condition),
				  timebin = seq(0, 17, length.out = 100),
				  voc_total = voc_values,
				  # sesion_id = first(data$session_id),
				  # age = c(21, 25, 30),
				  lp = levels(model_fits_time$fit_time_2$data$lp))

epreds <- add_epred_draws(newdata = nd,
						  object = model_fits_time$fit_time_2,
						  ndraws = 20,
						  re_formula = NA,
						  value = ".value") |> 
	mean_qi() |> 
	mutate(lp = factor(lp, 
					   levels = c("Monolingual (English)",
					   		   "Monolingual",
					   		   "Bilingual")),
		   voc_total = factor(x = voc_total, 
		   				   levels = voc_values,
		   				   labels = scales::percent(voc_values)))

epreds |> 
	ggplot(aes(timebin, .value,
			   colour = condition,
			   fill = condition)) +
	facet_grid(lp~voc_total) +
	geom_hline(yintercept = 0,
			   linetype = "dashed") +
	geom_ribbon(aes(ymin = .lower, ymax = .upper),
				alpha = 0.5,
				linewidth = 0) +
	geom_line(linewidth = 3/4) +
	labs(x = "Time (ms)",
		 y = "Logit (Target looking)",
		 colour = "Condition",
		 fill = "Condition") +
	scale_x_continuous(breaks = seq(0, 20, 4),
					   labels = \(x) format(x*100+200, big.mark = ",")) +
	theme(legend.position = "top",
		  panel.grid.major = element_line(colour = "grey",
		  								linetype = "dotted"))
```
